---
title: "wrangler"
author: "James Hamilton"
date: "March 2, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data Wrangling Exercise 1  
```{r echo=FALSE, message=FALSE}
library(dplyr)
library(tidyr)
```

## 0 - Load the data
Simply use read.csv
`r mydata <- read.csv("refine_original.csv", header = TRUE)`

## 1 - Clean up brand names
Using a combination of regex and tolower
```{r}
mydata$company <- tolower(mydata$company) %>%
  gsub("0", "o", .) %>%
  sub("k z", "kz", .) %>%
  sub("lv", "lev", .) %>%
  sub("^([fp].+)", "philips", .)
```

## 2 - Separate product code & number
Using tidyr separate function
```{r}
newcols = c("product_code","product_number")
mydata <- mydata %>%
  separate(., "Product.code...number", 
           newcols, sep = "-", remove = FALSE)
```

## 3 - Add product category
Using dplyr left_join and a mapping dataframe
```{r warning=FALSE}
codemapvector <- c("p","Smartphone","v","TV","x","Laptop","q","Tablet")
codemap <- data.frame(matrix(codemapvector, nrow=4, byrow=TRUE))
names(codemap) <- c("product_code", "product_category")
mydata <- left_join(mydata, codemap, by = "product_code")
```

## 4 - Add full addr. to allow geocoding
With dplyr mutate.
```{r}
mydata <- mydata %>% 
  mutate(., full_address = paste(address, city, country, sep = ", "))
```

## 5 - Create dummy vars: company & product category
Using mutate again.
```{r}
mydata <- mydata %>%
  mutate(., company_philips = mydata$company == "philips") %>%
  mutate(., company_akzo = mydata$company == "akzo") %>%
  mutate(., company_van_houten = mydata$company == "van houten") %>%
  mutate(., company_unilever = mydata$company == "unilever")

mydata <- mydata %>%
  mutate(., product_smartphone = mydata$product_code == "p") %>%
  mutate(., product_tv = mydata$product_code == "v") %>%
  mutate(., product_laptop = mydata$product_code == "x") %>%
  mutate(., product_tablet = mydata$product_code == "q")
```

## 6 - Write tidied data to refine_clean.csv
Using write.csv
```{r}
head(mydata)
write.csv(mydata, "refine_clean.csv", row.names = FALSE)
```